---
title: 'OHIBC: Calculate FIS scores'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 8, fig.height = 6, echo = TRUE, message = FALSE, warning = FALSE, fig.path = 'figs/')

dir_git <- '~/github/ohibc'
source(file.path(dir_git, 'src/R/common.R'))  ### an OHIBC specific version of common.R
dir_rgn <- file.path(dir_git, 'prep/regions')  ### github: general buffer region shapefiles
dir_anx <- file.path(dir_M, 'git-annex/bcprep')


### support scripts
#source(file.path(dir_git, 'src/R/rast_tools.R')) 
source("~/github/ohibc/calc_ohibc/calc_scores_fxns.R")

```

# Summary

This script takes the output from `functions.R` FIS model and plots individual stock scores by year and relative contribution to score.

Here we are calculating the **Base Model** before we run any scenarios. This model has the following characteristics:

- No catch weighting  
- No penalty for unassessed catch  
- Keep underfishing penalty  

------

```{r source fisheries function}

source("fis_function.R")

status <- FIS(layers) %>%
  left_join(get_rgn_names(), by = c('region_id' = 'rgn_id'))

```


## Stock scores and overall scores by region

``` {r assemble_dataframe}

## plotting fishery catch weighting by region
stock_plot_df <- read_csv('output/stock_scores.csv') %>%
  select(-X1) %>%
  filter(year > 1995) %>% #just to reduce the amount of data we are looking at
  left_join(get_rgn_names(), by = c('region_id' = 'rgn_id')) %>%
  group_by(region_id, year) %>%
  mutate(avg = mean(score, na.rm=T))

```

``` {r plot_fis_stock_scores}

rgn_avg <- stock_plot_df %>%
  select(region_id, year, avg, rgn_name) %>%
  rename(score = avg) %>%
  distinct()

ggplot(stock_plot_df, aes(x = year, y = score)) +
    ggtheme_plot() +
    geom_line(aes(group = stock_id, color = stock_id),
              lineend = 'round', alpha = 0.8) +
    labs(color = 'Stock ID',
         y     = 'Stock Score') +
    guides(colour = guide_legend(override.aes = list(size = 3)),
           size = 'none') +
    facet_wrap(~rgn_name) +
    geom_line(data = rgn_avg, aes(x = year, y = score, group = rgn_name), color = "black", lwd = 1)

```

#Fisheries status scores

``` {r plot_fis_status_scores}

ggplot(status, aes(x = year, y = score)) +
    ggtheme_plot() +
    geom_line(aes(group = rgn_name, color = rgn_name), alpha = 0.8) +
    labs(color = 'Region',
         y     = 'Fisheries Status') +
    ggtitle("Fisheries status in British Columbia")

```
