---
title: 'OHIBC: explore DFO data'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, message = FALSE, warning = FALSE)

dir_git     <- path.expand('~/github/ohibc')
source(file.path(dir_git, 'src/R/common.R'))  
dir_spatial <- file.path(dir_git, 'prep/_spatial')  
dir_anx     <- file.path(dir_M, 'git-annex/bcprep')
```

# Summary

Karen sent shapefiles for all stocks and years aggregated to OHI BC regions. This script just checks out that data.

```{r}
library(tidyverse)
library(sf)
```
Let's take a look at one file to see the data format

```{r}
halibut <- st_read(file.path(dir_anx, "_raw_data/dfo_khunter/fisheries/d2018/halibut/Halibut_2015.shp"), quiet = T)

#data held in the shapefile
DT::datatable(halibut)
```

Do they all look like this? Let's look at shrimp

```{r}
shrimp <- st_read(file.path(dir_anx, "_raw_data/dfo_khunter/fisheries/d2018/shrimp/Shrimp_2009.shp"), quiet = T)

#data held in the shapefile
DT::datatable(shrimp)
```

OH boy, let's look at more

```{r}
crab <- st_read(file.path(dir_anx, "_raw_data/dfo_khunter/fisheries/d2018/dungeness_crab/DCrab_2010.shp"), quiet = T)

#data held in the shapefile
DT::datatable(crab)
```

I emailed Karen at DFO to get the details. Here is the important information:

Column to use for all, if available is `sum_landed` which has Kilograms as units. 

>For shrimp:   
> "You need to sum columns _pinks thru spny for shrimp catch"
>
>For Dungeness:  
> "Use sum calc_wg for crab catch"
>
>Column to use for **Spot prawn** is `sum_prwn_w`. The units are in kilograms.
>
>Column to use for **Herring** is `Sum_Catch`. The units are in MT.
>
>**Boccacio** – I do not understand why this is not `Sum_Landed` (KGS) like the other groundfish. I’ll take a look at this. Could be an artifact of bycatch records in the DB.  Looks like there is a `sum_catch` in the output. Sorry about that. I was going quickly and didn’t catch this strange output.
>
>Column to use for **Geoduck** is `Tl_landed`. The units are in Pounds (LB).  

Created a function, `read_files` to read in each annual file for each species, aggregate, and make a dataframe with the columns we are interested in (region and catch).
```{r}
spp_folders <- list.files(file.path(dir_anx, '_raw_data/dfo_khunter/fisheries/d2018'))

read_files <- function(folder, column){
  
  files <- list.files(paste0(dir_anx, '/_raw_data/dfo_khunter/fisheries/d2018/', folder), full.names = T)
  
  out <- data.frame()
  
  for(yr in 2007:2015){
    #get the shapefile for yr
    yr_files <- files[str_detect(files, as.character(yr))]
    
    yr_shp_file <- yr_files[str_detect(yr_files, as.character(".shp"))]
    
    yr_shp <- st_read(yr_shp_file, quiet = T) %>%
      as.data.frame() %>%
      select(rgn_name, rgn_id, rgn_code, column) %>%
      mutate(year = yr)
    
    out <- rbind(yr_shp, out)
  }
    return(out)
    
}
```

Now we can get a single dataset for each stock by running the `read_files` function. 
```{r}

#arrowtooth flounder
af <- read_files("arrowtooth_flounder", "Sum_LANDED") %>%
  mutate(species = "arrowtooth_flounder",
         units   = "kg") %>%
  rename(harvest = "Sum_LANDED")
#boccacio
boc <- read_files("boccacio", "Sum_CATCH_") %>%
  mutate(species = "boccacio",
         units   = "kg") %>%
  rename(harvest = "Sum_CATCH_")
#dogfish
dog <- read_files("dogfish", "Sum_LANDED") %>%
  mutate(species = "dogfish",
         units   = "kg") %>%
  rename(harvest = "Sum_LANDED")
#dungness crab
crab <- read_files("dungeness_crab", "Sum_clc_wg") %>%
  mutate(species = "dungeness_crab",
         units   = "kg") %>%
  rename(harvest = "Sum_clc_wg")
#halibut
hal <- read_files("halibut", "Sum_LANDED") %>%
  mutate(species = "halibut",
         units   = "kg") %>%
  rename(harvest = "Sum_LANDED")
#herring
herr <- read_files("herring", "Sum_Catch") %>%
  mutate(species = "herring",
         units   = "metric_tons") %>%
  rename(harvest = "Sum_Catch")
#geoduck
geo <- read_files("geoduck", "Sum_Tt_lnd") %>%
  mutate(species = "geoduck",
         units   = "lbs") %>%
  rename(harvest = "Sum_Tt_lnd")
#lingcod
ling <- read_files("lingcod", "Sum_LANDED") %>%
  mutate(species = "lingcod",
         units   = "kg") %>%
  rename(harvest = "Sum_LANDED")
#rockfish
rfish <- read_files("rockfish", "Sum_LANDED") %>%
  mutate(species = "rockfish",
         units   = "kg") %>%
  rename(harvest = "Sum_LANDED")
#sablefish
sable <- read_files("sablefish", "Sum_LANDED") %>%
  mutate(species = "sablefish",
         units   = "kg") %>%
  rename(harvest = "Sum_LANDED")
#skates
skates <- read_files("skates", "Sum_LANDED") %>%
  mutate(species = "skates",
         units   = "kg") %>%
  rename(harvest = "Sum_LANDED")
#spot prawn
prawn <- read_files("spot_prawn", "Sum_prwn_w") %>%
  mutate(species = "spot_prawn",
         units   = "kg") %>%
  rename(harvest = "Sum_prwn_w")
#walleye pollock
pollock <- read_files("walleye_pollock", "Sum_LANDED") %>%
  mutate(species = "walleye_pollock",
         units   = "kg") %>%
  rename(harvest = "Sum_LANDED")
#soles (without southern rock sole)
soles <- read_files("soles", "Sum_LANDED") %>%
  mutate(species = "sole",
         units   = "kg") %>%
  rename(harvest = "Sum_LANDED")
#southern rock sole
srsole <- read_files("southernrocksole", "Sum_LANDED") %>%
  mutate(species = "srsole",
         units   = "kg") %>%
  rename(harvest = "Sum_LANDED")
```

Shrimp are trickier - we need to add across columns. So I'm doing this outside the `read_files` function. 

Karen says:  
>"You need to sum columns _pinks thru spny for shrimp catch"

```{r shrimp}

files <- list.files(paste0(dir_anx, '/_raw_data/dfo_khunter/fisheries/d2018/shrimp'), full.names = T)
  
  out <- data.frame()
  
  for(yr in 2007:2015){
    #get the shapefile for yr
    yr_files <- files[str_detect(files, as.character(yr))]
    
    yr_shp_file <- yr_files[str_detect(yr_files, as.character(".shp"))]
    
    yr_shp <- st_read(yr_shp_file, quiet = T) %>%
      as.data.frame() %>%
      mutate(harvest = Sum_pinks + Sum_sdstrp + Sum_prawns + Sum_hmpbck + Sum_docks + Sum_smth_p + Sum_flxd_p +Sum_othr_w + Sum_spny_p) %>%
      select(rgn_name, rgn_id, rgn_code, harvest) %>%
      mutate(year = yr)
    
    out <- rbind(yr_shp, out)
  }
  
shrimp <- out %>%
  mutate(species = "shrimp",
         units   = "kg")
```

Now I'm combining all dataframes into one, converting tons and lbs to kilograms, and adding in stock_id to match with RAM.

For Southern Rock sole, I used the map in [this assessment](http://waves-vagues.dfo-mpo.gc.ca/Library/363948.pdf), page 30, to make the delineations between regions.

```{r}
species_catch <- af %>%
  rbind(boc, dog, crab, hal, herr, geo, ling, rfish, sable, skates, prawn, pollock, soles, srsole, shrimp) %>%
  mutate(harvest_kgs = ifelse(units == "lbs", harvest/0.453592,  #pounds to kg
                              ifelse(units == "metric_tons", harvest*1000, harvest))) %>%
  select(-units) %>%
  mutate(stock_id = 
           case_when(
             species == "boccacio" ~ "BOCACCBCW",
             species == "lingcod" & rgn_name == "Strait of Georgia"  ~ "LINGCODSOG",
             species == "halibut" ~ "PHALBPAC",
             species == "sablefish" ~ "SABLEFPCAN",
             species == "srsole" & rgn_name == "North Coast" ~ "RSOLEHSTR", #rock sole in the North Coast is Hecate Strait stock
             species == "srsole" & rgn_name == "Aristazabal Island" ~ "RSOLEHSTR",
             species == "srsole" & rgn_name == "North Vancouver Island" ~ "RSOLE5AB",
             species == "srsole" & rgn_name == "Central Coast" ~ "RSOLE5AB",
             species == "srsole" & rgn_name == "Haida Gwaii" ~ "RSOLEHSTR"
             ))

write_csv(species_catch, "output/dfo_catch.csv")
```

I'm going to plot each species catch over time. Sent to Karen to look at. The Geoduck trends look questionable.

```{r plot, fig.width = 8, fig.height = 8}

ggplot(species_catch, aes(x = year, y = harvest_kgs, color = rgn_name)) +
  geom_line() +
  facet_wrap(~species, scales = 'free')

```













